### 2016-08-21 paper.tex
Initial commit for TPCP ACM CSUR Paper

% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

\usepackage[english]{babel}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage[pass]{geometry}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{url}
\usepackage{color}
\definecolor{nodecolor}{RGB}{202,202,202}

\widowpenalty=10000
\clubpenalty=10000

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{0000001.0000001}

%ISSN
\issn{1234-56789}

% Document starts
\begin{document}

% Page heads
\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}

% Title portion
\title{Total Platform Cyber Protection}
\author{GANG ZHOU
\affil{College of William and Mary}
YAFENG WU
\affil{University of Virginia}
TING YAN
\affil{Eaton Innovation Center}
TIAN HE
\affil{University of Minnesota}
CHENGDU HUANG
\affil{Google}
JOHN A. STANKOVIC
\affil{University of Virginia}
TAREK F. ABDELZAHER
\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
Multifrequency media access control has been well understood in
general wireless ad hoc networks, while in wireless sensor networks,
researchers still focus on single frequency solutions. In wireless
sensor networks, each device is typically equipped with a single
radio transceiver and applications adopt much smaller packet sizes
compared to those in general wireless ad hoc networks. Hence, the
multifrequency MAC protocols proposed for general wireless ad hoc
networks are not suitable for wireless sensor network applications,
which we further demonstrate through our simulation experiments. In
this article, we propose MMSN, which takes advantage of
multifrequency availability while, at the same time, takes into
consideration the restrictions of wireless sensor networks. Through
extensive experiments, MMSN exhibits the prominent ability to utilize
parallel transmissions among neighboring nodes. When multiple physical
frequencies are available, it also achieves increased energy
efficiency, demonstrating the ability to work against radio
interference and the tolerance to a wide range of measured time
synchronization errors.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
% End generated code
%

% We no longer use \terms command
%\terms{Design, Algorithms, Performance}

\keywords{Wireless sensor networks, media access control,
multi-channel, radio interference, time synchronization}

\acmformat{Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic,
and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially
designed for  wireless sensor network applications.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
This work is supported by the National Science Foundation, under
grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.

Author's addresses: G. Zhou, Computer Science Department,
College of William and Mary; Y. Wu  {and} J. A. Stankovic,
Computer Science Department, University of Virginia; T. Yan,
Eaton Innovation Center; T. He, Computer Science Department,
University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
(Current address) NASA Ames Research Center, Moffett Field, California 94035.
\end{bottomstuff}

\maketitle

\section{Introduction}

As a new technology, Wireless Sensor Networks (WSNs) has a wide
range of applications [Culler 2001,Bahl 2002,Akyildiz 2001], including
environment monitoring, smart buildings, medical care, industrial and
military applications. Among them, a recent trend is to develop
commercial sensor networks that require pervasive sensing of both
environment and human beings, for example, assisted living
[Akyildiz 2002,Harvard 2001,CROSSBOW] and smart homes
[Harvard 2001,Adya 2001,CROSSBOW].
% quote
\begin{quote}
``For these applications, sensor devices are incorporated into human
cloths [Natarajan 2001,Zhou 2006,Bahl 2002,Adya 2001] for monitoring
health related information like EKG readings, fall detection, and voice recognition".
\end{quote}
While collecting all these multimedia information
[Akyildiz 2002] requires a high network throughput, off-the-shelf
sensor devices only provide very limited bandwidth in a single
channel: 19.2Kbps in MICA2 [Bahl 2002] and 250Kbps in MICAz.

In this article, we propose MMSN, abbreviation for Multifrequency
Media access control for wireless Sensor Networks. The main
contributions of this work can be summarized as follows.
% itemize
\begin{itemize}
\item To the best of our knowledge, the MMSN protocol is the first
multifrequency MAC protocol especially designed for WSNs, in which
each device is equipped with a single radio transceiver and
the MAC layer packet size is very small.
\item Instead of using pairwise RTS/CTS frequency negotiation
[Adya 2001,Culler 2001; Tzamaloukas 2001; Zhou 2006],
we propose lightweight frequency assignments, which are good choices
for many deployed comparatively static WSNs.
\item We develop new toggle transmission and snooping techniques to
enable a single radio transceiver in a sensor device to achieve
scalable performance, avoiding the nonscalable ``one
control channel + multiple data channels'' design [Natarajan 2001].
\end{itemize}

% Head 1
\section{MMSN Protocol}

% Head 2
\subsection{Frequency Assignment}

We propose a suboptimal distribution to be used by each node, which is
easy to compute and does not depend on the number of competing
nodes. A natural candidate is an increasing geometric sequence, in
which
% Numbered Equation
\begin{equation}
\label{eqn:01}
P(t)=\frac{b^{\frac{t+1}{T+1}}-b^{\frac{t}{T+1}}}{b-1},
\end{equation}
where $t=0,{\ldots}\,,T$, and $b$ is a number greater than $1$.

In our algorithm, we use the suboptimal approach for simplicity and
generality. We need to make the distribution of the selected back-off
time slice at each node conform to what is shown in Equation
(\ref{eqn:01}). It is implemented as follows: First, a random
variable $\alpha$ with a uniform distribution within the interval
$(0, 1)$ is generated on each node, then time slice $i$ is selected
according to the following equation:
% Unnumbered Equation
\[
i=\lfloor(T+1)\log_b[\alpha(b-1)+1]\rfloor.
\]
It can be easily proven that the distribution of $i$ conforms to Equation
(\ref{eqn:01}).

So protocols [Bahl 2002,Culler 2001,Zhou 2006,Adya 2001,Culler 2001;
Tzamaloukas-01; Akyildiz-01] that use RTS/CTS
controls\footnote{RTS/CTS controls are required to be implemented by
802.11-compliant devices. They can be used as an optional mechanism
to avoid Hidden Terminal Problems in the 802.11 standard and
protocols based on those similar to [Akyildiz 2001] and
[Adya 2001].} for frequency negotiation and reservation are not
suitable for WSN applications, even though they exhibit good
performance in general wireless ad hoc
networks.

% Head 3
\subsubsection{Exclusive Frequency Assignment}

In exclusive frequency assignment, nodes first exchange their IDs
among two communication hops so that each node knows its two-hop
neighbors' IDs. In the second broadcast, each node beacons all
neighbors' IDs it has collected during the first broadcast period.

% Head 4
\paragraph{Eavesdropping}

Even though the even selection scheme leads to even sharing of
available frequencies among any two-hop neighborhood, it involves a
number of two-hop broadcasts. To reduce the communication cost, we
propose a lightweight eavesdropping scheme.

\subsection{Basic Notations}

As Algorithm~\ref{alg:one} states, for each frequency
number, each node calculates a random number (${\textit{Rnd}}_{\alpha}$) for
itself and a random number (${\textit{Rnd}}_{\beta}$) for each of its two-hop
neighbors with the same pseudorandom number generator.
% Algorithm
\begin{algorithm}[t]
\SetAlgoNoLine
\KwIn{Node $\alpha$'s ID ($ID_{\alpha}$), and node $\alpha$'s
neighbors' IDs within two communication hops.}
\KwOut{The frequency number ($FreNum_{\alpha}$) node $\alpha$ gets assigned.}
$index$ = 0; $FreNum_{\alpha}$ = -1\;
\Repeat{$FreNum_{\alpha} > -1$}{
        $Rnd_{\alpha}$ = Random($ID_{\alpha}$, $index$)\;
        $Found$ = $TRUE$\;
        \For{each node $\beta$ in $\alpha$'s two communication hops
    }{
      $Rnd_{\beta}$ = Random($ID_{\beta}$, $index$)\;
      \If{($Rnd_{\alpha} < Rnd_{\beta}$) \text{or} ($Rnd_{\alpha}$ ==
          $Rnd_{\beta}$ \text{and} $ID_{\alpha} < ID_{\beta}$)\;
      }{
        $Found$ = $FALSE$; break\;
      }
        }
     \eIf{$Found$}{
           $FreNum_{\alpha}$ = $index$\;
         }{
           $index$ ++\;
     }
      }
\caption{Frequency Number Computation}
\label{alg:one}
\end{algorithm}

Bus masters are divided into two disjoint sets, $\mathcal{M}_{RT}$
and $\mathcal{M}_{NRT}$.
% description
\begin{description}
\item[RT Masters]
$\mathcal{M}_{RT}=\{ \vec{m}_{1},\dots,\vec{m}_{n}\}$ denotes the
$n$ RT masters issuing real-time constrained requests. To model the
current request issued by an $\vec{m}_{i}$ in $\mathcal{M}_{RT}$,
three parameters---the recurrence time $(r_i)$, the service cycle
$(c_i)$, and the relative deadline $(d_i)$---are used, with their
relationships.
\item[NRT Masters]
$\mathcal{M}_{NRT}=\{ \vec{m}_{n+1},\dots,\vec{m}_{n+m}\}$ is a set
of $m$ masters issuing nonreal-time constrained requests. In our
model, each $\vec{m}_{j}$ in $\mathcal{M}_{NRT}$ needs only one
parameter, the service cycle, to model the current request it
issues.
\end{description}

Here, a question may arise, since each node has a global ID. Why
don't we just map nodes' IDs within two hops into a group of
frequency numbers and assign those numbers to all nodes within two
hops?

\section{Simulator}
\label{sec:sim}

If the model checker requests successors of a state which are not
created yet, the state space uses the simulator to create the
successors on-the-fly. To create successor states the simulator
conducts the following steps.
% enumerate
\begin{enumerate}
\item Load state into microcontroller model.
\item Determine assignments needed for resolving nondeterminism.
\item For each assignment.
      \begin{enumerate}
      \item either call interrupt handler or simulate effect of next instruction, or
      \item evaluate truth values of atomic propositions.
      \end{enumerate}
\item Return resulting states.
\end{enumerate}
Figure~\ref{fig:one} shows a typical microcontroller C program that
controls an automotive power window lift. The program is one of the
programs used in the case study described in Section~\ref{sec:sim}.
At first sight, the programs looks like an ANSI~C program. It
contains function calls, assignments, if clauses, and while loops.
% Figure
\begin{figure}
\caption{Code before preprocessing.}
\label{fig:one}
\end{figure}

\subsection{Problem Formulation}

The objective of variable coalescence-based offset assignment is to find
both the coalescence scheme and the MWPC on the coalesced graph. We start
with a few definitions and lemmas for variable coalescence.

% Enunciations
\begin{definition}[Coalesced Node (C-Node)]A C-node is a set of
live ranges (webs) in the AG or IG that are coalesced. Nodes within the same
C-node cannot interfere with each other on the IG. Before any coalescing is
done, each live range is a C-node by itself.
\end{definition}

\begin{definition}[C-AG (Coalesced Access Graph)]The C-AG is the access
graph after node coalescence, which is composed of all C-nodes and C-edges.
\end{definition}

\begin{lemma}
The C-MWPC problem is NP-complete.
\end{lemma}
\begin{proof} C-MWPC can be easily reduced to the MWPC problem assuming a
coalescence graph without any edge or a fully connected interference graph.
Therefore, each C-node is an uncoalesced live range after value separation
and C-PC is equivalent to PC. A fully connected interference graph is made
possible when all live ranges interfere with each other. Thus, the C-MWPC
problem is NP-complete.
\end{proof}

\begin{lemma}[Lemma Subhead]The solution to the C-MWPC problem is no
worse than the solution to the MWPC.
\end{lemma}
\begin{proof}
Simply, any solution to the MWPC is also a solution to the
C-MWPC. But some solutions to C-MWPC may not apply to the MWPC (if any
coalescing were made).
\end{proof}

\section{Performance Evaluation}

During all the experiments, the Geographic Forwarding (GF)
[Akyildiz 2001] routing protocol is used. GF exploits geographic
information of nodes and conducts local data-forwarding to achieve
end-to-end routing. Our simulation is
configured according to the settings in
Table~\ref{tab:one}. Each run lasts for 2 minutes and
repeated 100 times. For each data value we present in the results,
we also give its 90\% confidence interval.
% Table
\begin{table}%
\tbl{Simulation Configuration\label{tab:one}}{%
\begin{tabular}{|l|l|}
\hline
TERRAIN{$^a$}   & (200m$\times$200m) Square\\\hline
Node Number     & 289\\\hline
Node Placement  & Uniform\\\hline
Application     & Many-to-Many/Gossip CBR Streams\\\hline
Payload Size    & 32 bytes\\\hline
Routing Layer   & GF\\\hline
MAC Layer       & CSMA/MMSN\\\hline
Radio Layer     & RADIO-ACCNOISE\\\hline
Radio Bandwidth & 250Kbps\\\hline
Radio Range     & 20m--45m\\\hline
\end{tabular}}
\begin{tabnote}%
\Note{Source:}{This is a table
sourcenote. This is a table sourcenote. This is a table
sourcenote.}
\vskip2pt
\Note{Note:}{This is a table footnote.}
\tabnoteentry{$^a$}{This is a table footnote. This is a
table footnote. This is a table footnote.}
\end{tabnote}%
\end{table}%

\section{Conclusions}

In this article, we develop the first multifrequency MAC protocol for
WSN applications in which each device adopts a
single radio transceiver. The different MAC design requirements for
WSNs and general wireless ad-hoc networks are
compared, and a complete WSN multifrequency MAC design (MMSN) is
put forth. During the MMSN design, we analyze and evaluate different
choices for frequency assignments and also discuss the nonuniform
back-off algorithms for the slotted media access design.

% Start of "Sample References" section

\section{Typical references in new ACM Reference Format}
A paginated journal article \cite{Abril07}, an enumerated
journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
followed by the same example, however we only output the series if the volume number is given
\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
an article in a proceedings (of a conference, symposium, workshop for example)
(paginated proceedings article) \cite{Andler79}, a proceedings article
with all possible elements \cite{Smith10}, an example of an enumerated
proceedings article \cite{VanGundy07},
an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
a master's thesis: \cite{anisi03}, an online document / world wide web resource \cite{Thornburg01}, \cite{Ablamowicz07},
\cite{Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
Boris / Barbara Beeton: multi-volume works as books
\cite{MR781536} and \cite{MR781537}.

% Appendix
\appendix
\section*{APPENDIX}
\setcounter{section}{1}
In this appendix, we measure
the channel switching time of Micaz [CROSSBOW] sensor devices.
In our experiments, one mote alternatingly switches between Channels
11 and 12. Every time after the node switches to a channel, it sends
out a packet immediately and then changes to a new channel as soon
as the transmission is finished. We measure the
number of packets the test mote can send in 10 seconds, denoted as
$N_{1}$. In contrast, we also measure the same value of the test
mote without switching channels, denoted as $N_{2}$. We calculate
the channel-switching time $s$ as
\begin{eqnarray}%
s=\frac{10}{N_{1}}-\frac{10}{N_{2}}. \nonumber
\end{eqnarray}%
By repeating the experiments 100 times, we get the average
channel-switching time of Micaz motes: 24.3$\mu$s.

\appendixhead{ZHOU}

% Acknowledgments
\begin{acks}
The authors would like to thank Dr. Maura Turolla of Telecom
Italia for providing specifications about the application scenario.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{acmsmall-sample-bibfile}
\bibliography{papers}
                             % Sample .bib file with references that match those in
                             % the 'Specifications Document (V1.5)' as well containing
                             % 'legacy' bibs and bibs with 'alternate codings'.
                             % Gerry Murray - March 2012

% History dates
\received{February 2007}{March 2009}{June 2009}

\end{document}
% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM

### 2016-08-21 paper.tex
added tags for ccs concepts, used ccs code generator, might need more codes custom modded the numbers to 400, to emphasize that all are same level of importance default is 500, 300, 100

@@
 \title{Total Platform Cyber Protection}
-\author{GANG ZHOU
-\affil{College of William and Mary}
-YAFENG WU
-\affil{University of Virginia}
-TING YAN
-\affil{Eaton Innovation Center}
-TIAN HE
-\affil{University of Minnesota}
-CHENGDU HUANG
-\affil{Google}
-JOHN A. STANKOVIC
-\affil{University of Virginia}
-TAREK F. ABDELZAHER
-\affil{University of Illinois at Urbana-Champaign}}
+
+%TODO Adam: Not sure if anon submission so just leaving like this for now
+\author{
+%First Author
+%\affil{Arizona State University}
+%Second Author
+%\affil{Arizona State University}
+%Third Author
+%\affil{Arizona State University}
+%Fourth Author
+%\affil{Arizona State University}
+%Fifth Author
+%\affil{Arizona State University}
+%Sixth Author
+%\affil{Arizona State University}
+%Seventh Author
+%\affil{Arizona State University}
+}
 % NOTE! Affiliations placed here should be for the institution where the
@@
 
+% TODO Sai: Discuss story with yeganeh and get a nice little abstract here
 \begin{abstract}
@@
 	<concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
+	<concept_id>10002978.10003001.10010777</concept_id>
+	<concept_desc>Security and privacy~Hardware attacks and countermeasures</concept_desc>
+	<concept_significance>400</concept_significance>
 	</concept>
 	<concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
+	<concept_id>10002978.10003001.10003003</concept_id>
+	<concept_desc>Security and privacy~Embedded systems security</concept_desc>
+	<concept_significance>400</concept_significance>
 	</concept>
 	<concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
+	<concept_id>10002978.10003006.10003007</concept_id>
+	<concept_desc>Security and privacy~Operating systems security</concept_desc>
+	<concept_significance>400</concept_significance>
 	</concept>
 	<concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+	<concept_id>10002978.10003022</concept_id>
+	<concept_desc>Security and privacy~Software and application security</concept_desc>
+	<concept_significance>400</concept_significance>
 	</concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
+\ccsdesc[400]{Security and privacy~Hardware attacks and countermeasures}
+\ccsdesc[400]{Security and privacy~Embedded systems security}
+\ccsdesc[400]{Security and privacy~Operating systems security}
+\ccsdesc[400]{Security and privacy~Software and application security}
 %
@@
 
-\keywords{Wireless sensor networks, media access control,
-multi-channel, radio interference, time synchronization}
+%TODO Sai: Add keywords here
+\keywords{}
 

### 2016-08-21 paper.tex
added in the intro and bg sections

@@
 \definecolor{nodecolor}{RGB}{202,202,202}
+\newlength\framesep
+\setlength\framesep{16pt}
+\pgfdeclarelayer{background}
+\pgfsetlayers{background,main}
+\pgfkeys{
+	/tikz/node distance/.append code={
+		\pgfkeyssetvalue{/tikz/node distance value}{#1}
+	}
+}
 
@@
 \section{Introduction}
+Traditional security defenses for the computing environment have
+focused on securing the \emph{border} of the network, as this is the
+entry place for most attackers. However, even with sophisticated
+border technologies in place, there are constant attacks and data
+breaches against our networks. The goal of this work is to survey the
+state of security for the entire computing platform, in order to
+identify specific areas that are underserved by the research community
+and that require additional research and investment.
 
-As a new technology, Wireless Sensor Networks (WSNs) has a wide
-range of applications [Culler 2001,Bahl 2002,Akyildiz 2001], including
-environment monitoring, smart buildings, medical care, industrial and
-military applications. Among them, a recent trend is to develop
-commercial sensor networks that require pervasive sensing of both
-environment and human beings, for example, assisted living
-[Akyildiz 2002,Harvard 2001,CROSSBOW] and smart homes
-[Harvard 2001,Adya 2001,CROSSBOW].
-% quote
-\begin{quote}
-``For these applications, sensor devices are incorporated into human
-cloths [Natarajan 2001,Zhou 2006,Bahl 2002,Adya 2001] for monitoring
-health related information like EKG readings, fall detection, and voice recognition".
-\end{quote}
-While collecting all these multimedia information
-[Akyildiz 2002] requires a high network throughput, off-the-shelf
-sensor devices only provide very limited bandwidth in a single
-channel: 19.2Kbps in MICA2 [Bahl 2002] and 250Kbps in MICAz.
+Modern computing infrastructure serves a wide array of needs, with
+diverse technologies and implementations: embedded systems, cloud
+computing, sensor networks, desktops, mobile devices, and industrial
+control systems. Rather than consider each of these computing
+platforms independently, in this work we abstracted the computing
+platform into several layers, each of which are applicable to every
+computing infrastructure. In particular, we analyzed the security
+research performed at the following layers of the computing platform:
+hardware, firmware, bus, hypervisor, operating system, application,
+and network layers (Figure~1).
 
-In this article, we propose MMSN, abbreviation for Multifrequency
-Media access control for wireless Sensor Networks. The main
-contributions of this work can be summarized as follows.
-% itemize
-\begin{itemize}
-\item To the best of our knowledge, the MMSN protocol is the first
-multifrequency MAC protocol especially designed for WSNs, in which
-each device is equipped with a single radio transceiver and
-the MAC layer packet size is very small.
-\item Instead of using pairwise RTS/CTS frequency negotiation
-[Adya 2001,Culler 2001; Tzamaloukas 2001; Zhou 2006],
-we propose lightweight frequency assignments, which are good choices
-for many deployed comparatively static WSNs.
-\item We develop new toggle transmission and snooping techniques to
-enable a single radio transceiver in a sensor device to achieve
-scalable performance, avoiding the nonscalable ``one
-control channel + multiple data channels'' design [Natarajan 2001].
-\end{itemize}
+The goal of this work is to focus research effort on \emph{securing
+	the entire computing platform.} An attack must effectively target a
+specific vulnerability in a specific layer of the computing stack, and
+an attacker uses that vulnerability to establish persistence on the
+machine, potentially attacking the underlying computing layers of the
+same machine or leveraging their place in the network to attack other
+machines. Therefore, if we wish to increase the security of our
+computing systems, and reduce the number and scope of security
+breaches, it is essential that we encourage focus on novel ideas,
+algorithms, and techniques to secure every level of the computing
+stack.
+
+In this report, we will first discuss the computing stack, and, for
+every layer, we will describe the layer, identify primary threats
+against the layer, and summarize the research on attempts to secure
+that layer, and finally we will determine areas that require further
+research investment to ensure the security of the layer.\input{fig_layers}
+
+\section{Background}
+Here, we consider the diverse array of computing platforms as a number
+of abstracted layers: hardware, firmware, bus, hypervisor, operating
+system, application, and network layers. In this way, we can
+categorize the research efforts that target each layer.
+
+The layers are arranged from lowest to highest layer. The lower layer
+has more control over the computing system, while the upper layers
+have less control. For instance, the Hardware layer has total control
+over the physical hardware and memory of the machine, while an
+application running in the Application layer can only see a portion of
+the machine's memory (because of the virtual memory used by the
+Operating System).
+
+A typical compromise, in relation to the proposed layers, is the
+following: A vulnerability in an application is targeted, thus
+allowing the attacker full control of the application layer. From
+here, the attacker can scan the other applications on the network and
+attempt to exploit those applications, thus spreading horizontally
+throughout the network. Alternatively, as the application layer is the
+least privileged and least persistent of the layers, the attack can
+attempt to further the compromise on the lower layers of the machine.
+For instance, the attacker can exploit the hypervisor layer to
+compromise the host hypervisor. From this vantage point, the attacker
+can transparently extract information from all the other guest
+Operating Systems. More insidiously, the attacker can even exploit the
+firmware level, which allows the attacker to persist even after
+a complete reinstall of the operating system.
+
+In our abstract model, the lower levels have more control over the
+computing system, yet they have a smaller attacker surface. However,
+the lower levels can be attacked by exploiting the upper levels. Also, the upper
+levels trust the abstractions provided by the lower levels. For
+instance, the application-layer visualization of an industrial control
+system trusts that the data reported by the hardware layer is
+accurate. If an attacker is able to control and manipulate the
+hardware to falsely report the state of the system, the entire
+integrity of the system can be compromised.
+
+The interconnected nature of the layers means that the security of the
+entire layer must be a priority. However, at the same time, the
+security improvement of any layer will have an amplifying effect on
+the security of the whole system. For instance, improvement to the
+security of the application layer decreases the chance of an
+application compromise, which further decreases the risk of the attack
+propagating throughout the network and through the other layers.
 

### 2016-08-21 paper.tex
added hardware and firmware sections, made biblio work

@@
 
-% Head 1
-\section{MMSN Protocol}
+\section{Hardware}
 
-% Head 2
-\subsection{Frequency Assignment}
+Hardware underpins all of modern computing. While most software is
+constructed and written in such a way as to abstract the
+peculiarities of the underlying hardware, the software cannot
+execute without hardware. In our abstracted model of the computing
+stack, hardware is at the bottom, which means that it has complete
+control of the computing system.
 
-We propose a suboptimal distribution to be used by each node, which is
-easy to compute and does not depend on the number of competing
-nodes. A natural candidate is an increasing geometric sequence, in
-which
-% Numbered Equation
-\begin{equation}
-\label{eqn:01}
-P(t)=\frac{b^{\frac{t+1}{T+1}}-b^{\frac{t}{T+1}}}{b-1},
-\end{equation}
-where $t=0,{\ldots}\,,T$, and $b$ is a number greater than $1$.
+For hardware here, we consider all types of hardware in the computing
+system: CPUs, SoCs, or any other integrated circuit (IC) that performs
+computation or communication on behalf of the upper computing layers.
+This also includes networking cards and other peripheral devices
+(Bluetooth or USB devices), which may have their own special purpose
+embedded hardware.
 
-In our algorithm, we use the suboptimal approach for simplicity and
-generality. We need to make the distribution of the selected back-off
-time slice at each node conform to what is shown in Equation
-(\ref{eqn:01}). It is implemented as follows: First, a random
-variable $\alpha$ with a uniform distribution within the interval
-$(0, 1)$ is generated on each node, then time slice $i$ is selected
-according to the following equation:
-% Unnumbered Equation
-\[
-i=\lfloor(T+1)\log_b[\alpha(b-1)+1]\rfloor.
-\]
-It can be easily proven that the distribution of $i$ conforms to Equation
-(\ref{eqn:01}).
+A key difficulty in analyzing the security concerns of hardware is
+reverse engineering the workings of the hardware. Traditional
+techniques to reverse engineer hardware require destroying the chip,
+and if a single hardware component is destroyed to reverse engineer,
+there is no guarantee that other hardware is identical to the reverse
+engineered chip.
 
-So protocols [Bahl 2002,Culler 2001,Zhou 2006,Adya 2001,Culler 2001;
-Tzamaloukas-01; Akyildiz-01] that use RTS/CTS
-controls\footnote{RTS/CTS controls are required to be implemented by
-802.11-compliant devices. They can be used as an optional mechanism
-to avoid Hidden Terminal Problems in the 802.11 standard and
-protocols based on those similar to [Akyildiz 2001] and
-[Adya 2001].} for frequency negotiation and reservation are not
-suitable for WSN applications, even though they exhibit good
-performance in general wireless ad hoc
-networks.
+\subsection{Threats}
 
-% Head 3
-\subsubsection{Exclusive Frequency Assignment}
+As the lower-most level of the computing stack, all layers implicitly
+trust that the hardware is executing correctly. Therefore, if an
+adversary is able to maliciously control the hardware execution, they
+can compromise all layers of the computing stack. Thus, the security
+of the hardware is of paramount importance.
 
-In exclusive frequency assignment, nodes first exchange their IDs
-among two communication hops so that each node knows its two-hop
-neighbors' IDs. In the second broadcast, each node beacons all
-neighbors' IDs it has collected during the first broadcast period.
+The main threat considered by the literature on hardware is the threat
+of Trojan logic. This is malicious logic that is introduced into the
+% adversary, by definition is a malicious entity.
+hardware chip by an adversary. Trojan introduction can be
+done at several points of the hardware lifecycle:
 
-% Head 4
-\paragraph{Eavesdropping}
+\begin{itemize}
 
-Even though the even selection scheme leads to even sharing of
-available frequencies among any two-hop neighborhood, it involves a
-number of two-hop broadcasts. To reduce the communication cost, we
-propose a lightweight eavesdropping scheme.
+	\item \textbf{Design}. A malicious insider can incorporate a Trojan into the
+	design of the hardware.
 
-\subsection{Basic Notations}
+	\item \textbf{Third-party IP}. Oftentimes, hardware designers will take
+	advantage of the functionality of third-party components, using them
+	as a black-box. Thus, a third-party can include malicious Trojan
+	logic.
 
-As Algorithm~\ref{alg:one} states, for each frequency
-number, each node calculates a random number (${\textit{Rnd}}_{\alpha}$) for
-itself and a random number (${\textit{Rnd}}_{\beta}$) for each of its two-hop
-neighbors with the same pseudorandom number generator.
-% Algorithm
-\begin{algorithm}[t]
-\SetAlgoNoLine
-\KwIn{Node $\alpha$'s ID ($ID_{\alpha}$), and node $\alpha$'s
-neighbors' IDs within two communication hops.}
-\KwOut{The frequency number ($FreNum_{\alpha}$) node $\alpha$ gets assigned.}
-$index$ = 0; $FreNum_{\alpha}$ = -1\;
-\Repeat{$FreNum_{\alpha} > -1$}{
-        $Rnd_{\alpha}$ = Random($ID_{\alpha}$, $index$)\;
-        $Found$ = $TRUE$\;
-        \For{each node $\beta$ in $\alpha$'s two communication hops
-    }{
-      $Rnd_{\beta}$ = Random($ID_{\beta}$, $index$)\;
-      \If{($Rnd_{\alpha} < Rnd_{\beta}$) \text{or} ($Rnd_{\alpha}$ ==
-          $Rnd_{\beta}$ \text{and} $ID_{\alpha} < ID_{\beta}$)\;
-      }{
-        $Found$ = $FALSE$; break\;
-      }
-        }
-     \eIf{$Found$}{
-           $FreNum_{\alpha}$ = $index$\;
-         }{
-           $index$ ++\;
-     }
-      }
-\caption{Frequency Number Computation}
-\label{alg:one}
-\end{algorithm}
+	\item \textbf{Compilation of Design}. When the design of the hardware is
+	compiled, but before it is sent for fabrication, the design could be
+	changed to incorporate the Trojan logic. Note that this differs from
+	Design-time, because the Trojan logic is not included in the design.
+	Consider the case of malware on a hardware designer's machine, which
+	changes the design when it is compiled.
 
-Bus masters are divided into two disjoint sets, $\mathcal{M}_{RT}$
-and $\mathcal{M}_{NRT}$.
-% description
-\begin{description}
-\item[RT Masters]
-$\mathcal{M}_{RT}=\{ \vec{m}_{1},\dots,\vec{m}_{n}\}$ denotes the
-$n$ RT masters issuing real-time constrained requests. To model the
-current request issued by an $\vec{m}_{i}$ in $\mathcal{M}_{RT}$,
-three parameters---the recurrence time $(r_i)$, the service cycle
-$(c_i)$, and the relative deadline $(d_i)$---are used, with their
-relationships.
-\item[NRT Masters]
-$\mathcal{M}_{NRT}=\{ \vec{m}_{n+1},\dots,\vec{m}_{n+m}\}$ is a set
-of $m$ masters issuing nonreal-time constrained requests. In our
-model, each $\vec{m}_{j}$ in $\mathcal{M}_{NRT}$ needs only one
-parameter, the service cycle, to model the current request it
-issues.
-\end{description}
+	\item \textbf{Manufacture}. A malicious manufacturer can introduce Trojan logic into hardware after the design is received but before the hardware
+	is fabricated.
 
-Here, a question may arise, since each node has a global ID. Why
-don't we just map nodes' IDs within two hops into a group of
-frequency numbers and assign those numbers to all nodes within two
-hops?
+	\item \textbf{Delivery}. A malicious hardware component containing Trojan logic can be swapped with a benign component after fabrication but before
+	delivery and installation.
 
-\section{Simulator}
-\label{sec:sim}
+\end{itemize}
 
-If the model checker requests successors of a state which are not
-created yet, the state space uses the simulator to create the
-successors on-the-fly. To create successor states the simulator
-conducts the following steps.
-% enumerate
-\begin{enumerate}
-\item Load state into microcontroller model.
-\item Determine assignments needed for resolving nondeterminism.
-\item For each assignment.
-      \begin{enumerate}
-      \item either call interrupt handler or simulate effect of next instruction, or
-      \item evaluate truth values of atomic propositions.
-      \end{enumerate}
-\item Return resulting states.
-\end{enumerate}
-Figure~\ref{fig:one} shows a typical microcontroller C program that
-controls an automotive power window lift. The program is one of the
-programs used in the case study described in Section~\ref{sec:sim}.
-At first sight, the programs looks like an ANSI~C program. It
-contains function calls, assignments, if clauses, and while loops.
-% Figure
-\begin{figure}
-\caption{Code before preprocessing.}
-\label{fig:one}
-\end{figure}
+The malicious Trojan logic behavior comes in three different forms:
+% TODO: Sai: is this three or two??
 
-\subsection{Problem Formulation}
+\begin{itemize}
 
-The objective of variable coalescence-based offset assignment is to find
-both the coalescence scheme and the MWPC on the coalesced graph. We start
-with a few definitions and lemmas for variable coalescence.
+	\item \textbf{Trigger-based}. The malicious behavior or logic is only expressed when certain conditions are met. Consider a Trojan logic in a CPU,
+	which changes the executing code to Ring 0 (with the most
+	privileges) when a specific sequence of instructions are issued. The
+	trigger can be internal to the chip, such as in the CPU case, or
+	could be external, for instance a certain radio frequency will
+	trigger the malicious code.
 
-% Enunciations
-\begin{definition}[Coalesced Node (C-Node)]A C-node is a set of
-live ranges (webs) in the AG or IG that are coalesced. Nodes within the same
-C-node cannot interfere with each other on the IG. Before any coalescing is
-done, each live range is a C-node by itself.
-\end{definition}
+	\item \textbf{Constant}. The malicious behavior or logic is continually
+	executed and subtly changes the output of the hardware in a way that
+	compromises security. Consider a Trojan logic of a Wi-Fi card, which
+	leaks the Wi-Fi passphrase by embedding it in the operating
+	parameters of the Wi-Fi radio.
 
-\begin{definition}[C-AG (Coalesced Access Graph)]The C-AG is the access
-graph after node coalescence, which is composed of all C-nodes and C-edges.
-\end{definition}
+\end{itemize}
 
-\begin{lemma}
-The C-MWPC problem is NP-complete.
-\end{lemma}
-\begin{proof} C-MWPC can be easily reduced to the MWPC problem assuming a
-coalescence graph without any edge or a fully connected interference graph.
-Therefore, each C-node is an uncoalesced live range after value separation
-and C-PC is equivalent to PC. A fully connected interference graph is made
-possible when all live ranges interfere with each other. Thus, the C-MWPC
-problem is NP-complete.
-\end{proof}
+These features of Trojan logic in hardware drive the research that
+attempts to detect them.
 
-\begin{lemma}[Lemma Subhead]The solution to the C-MWPC problem is no
-worse than the solution to the MWPC.
-\end{lemma}
-\begin{proof}
-Simply, any solution to the MWPC is also a solution to the
-C-MWPC. But some solutions to C-MWPC may not apply to the MWPC (if any
-coalescing were made).
-\end{proof}
+\subsection{Research}
+
+Most research into hardware Trojans deal with using side-channel
+analysis to identify the malicious Trojan behavior~\cite{lee2005,
+agrawal2007, banga2008, banga2008a, jin2008, du2010, narasimhan2011,
+narasimhan2012, cao2013, davoodi2013, dworak2013, forte2013,
+liu2013, liu2014, bao2015}. This research uses power analysis, heat
+analysis, or other side-effects to determine that the hardware is
+executing abnormal logic. Unfortunately, this typically requires
+building a model based on a ``golden'' or known-good hardware, which
+is often not accessible. Furthermore, as chips become more complex,
+the Trojan logic become an increasingly smaller portion of the chip,
+which in turn reduces the side-channel behavior of the chip
+significantly.
+
+Other research has demonstrated the attacks that are possible by
+actually implementing Trojan behavior~\cite{king2008, jin2010,
+sturton2011, liu2013, jin2014, zhang2014}. This line of research
+shows that Trojan logic is feasible for an attack.
+
+Of particular practical interest is a description of \emph{actual
+Trojan logic} discovered in the ProASIC3 Flash FPGA military
+chip~\cite{skorobogatov2012}. While in this case it was not proved
+that the logic introduced to the chip was intentionally malicious,
+nevertheless there existed functionality in the hardware that was
+explicitly denied by the hardware designer.
+
+Another avenue of research is to change the design of the chip itself,
+in order to improve the testability of the chip or to increase the
+effectiveness of the side-channel analysis~\cite{chakraborty2008,
+li2008, salmani2009, hicks2010, rosenfeld2011, waksman2011,
+bhunia2013, davoodi2013, rajendran2013, rolt2014}. This line of
+research attempts to rethink designing the hardware itself. However, a
+major drawback is that this can increase the cost of the hardware and
+it must be somehow mandated to be used in COTS components.
+
+JTAG is a main avenue of reverse engineering hardware, so some work
+has investigated the problem of securing JTAG access to
+hardware~\cite{rosenfeld2010, ren2015}. The main idea is to allow only
+authenticated users to have access to the JTAG functionality, however
+this increases the cost and complexity of implemented JTAG.
+
+Another approach is to change the design of the chip to support
+runtime monitoring~\cite{waksman2010}. In this way, the hardware
+is monitoring itself, 
+rather than external analysis of
+side-channels. However, this has all the same drawbacks of changing
+the design of hardware, in addition to not being robust in the face of
+a powerful attacker who is able to alter the design.
+
+Recent approaches suggest to use machine-learning techniques to
+identify the Trojan logic~\cite{haider2015}. Another approach attempts
+to use a formal verification technique in order to prove that no
+malicious logic was added to the hardware~\cite{guo2015}.
+
+Finally, there are several surveys conducted on hardware security,
+which provide a good overview of the field~\cite{wang2008,
+tehranipoor2010, guin2014, guin2014a, guin2014b}.
+
+Important problems include how to deal with counterfeit chips, better
+non-destructive examination methods, continuous monitoring of the
+integrity not just at boot time, and detection of backdoors and 
+``undocumented features''.  Hardware design can also be done to
+aid this process, e.g., having some part of the chip consume more power
+to make it more identifiable for authentication purposes.  Other
+possibilities include running multiple chips with different
+implementations side-by-side and comparing results to detect unexpected
+behavior.
+
+\subsection{Suggested Focus Areas}
+
+A key problem with hardware security is non-destructive IC
+examination, identification, and authentication - how to ensure that
+the chip that is designed is what is actually fabricated, with no
+additional logic.
+
+An additional neglected area is how to detect malicious Trojan logic
+in third-party components. This would require some type of static
+analysis or simulated dynamic analysis of third-party components,
+which could detect Trojan logic.
+
+Another approach that could be effective is how to properly test the
+functionality of hardware after it has been fabricated. This could
+include changing the design of the chip to allow better/more guided
+fuzz testing of the chip itself, in order to discover the Trojan
+logic.
+
+Finally, a prevention approach could be taken, to monitoring the upper
+layers of the computing stack for attempts to exploit the underlying
+hardware. This can also include changing or altering the firmware
+layer in such a way that it is not possible to exercise the Trojan
+logic, potentially using randomization techniques similar to ASLR.
+
+\section{Firmware}
+
+%Note: I did not remove the previous versions of those paragraphs
+%(commented them) so feel free to check the previous versions and 
+%revert them if my changes have problems.
+
+%Firmware is the layer of software that sits and interacts directly
+%with the hardware. Typically, the firmware is controlled from the
+%upper layers by device drivers in the Operating System.
+
+%Faris:
+Firmware is the layer of software that sits and interacts directly
+with the hardware. Typically, the firmware is controlled by the
+device drivers in the upper layers of the Operating System.
+
+%Many different hardware contain firmware, for instance the boot BIOS
+%on a motherboard is a form of firmware, embedded systems often contain
+%specific firmware, hardware peripherals such as network cards contain
+%firmware as well. We consider all such areas as firmware.
+
+%Faris:
+Many different hardware contain firmware. For instance: the boot BIOS
+on a motherboard, embedded systems that contain
+specific firmware, hardware peripherals such as network cards. 
+We consider all such areas as firmware.
+
+%The key traits of firmware is that, although it is binary code, it is
+%operating/interacting directly with the hardware, so there are no
+%%OS-style abstraction layers such as virtual memory, well-defined
+%system calls, or hardware abstraction. Therefore, each firmware is
+%binary code which is specifically tailored to a specific hardware
+%platform.
+
+%Faris:
+Although firmware is binary code, it is
+operating/interacting directly with the hardware, so there are no
+OS-style abstraction layers such as virtual memory, well-defined
+system calls, or hardware abstraction which is the key trait of firmware. 
+Therefore, each firmware is specifically tailored to a specific hardware platform.
+
+%Due to the hardware-specific nature of firmware, \emph{reverse
+%	engineering} is critical to understanding the firmware. In addition,
+%reverse engineering is particularly challenging because source code is
+%frequently unavailable, and vendors' desire to protect their
+%intellectual property makes them actively obfuscate their
+%firmware. All this means that, extracting semantically useful
+%information from firmware is a key first step to performing any
+%security analysis of the firmware.
+
+%Faris:
+Due to the hardware-specific nature of firmware, \emph{reverse
+	engineering} is critical to understand it. In addition,
+reverse engineering is particularly challenging because the source code is
+frequently unavailable, and vendors' desire to protect their
+intellectual property makes them actively obfuscate their
+firmware. This means, extracting semantically useful
+information from firmware is the first and key step to perform any
+security analysis of the firmware.
+
+\subsection{Threats}
+
+As the lowest-level of software code in the computing stack, threats
+to firmware can compromise the security of the entire machine.
+Firmware security is concerned with two main threats: (1)
+vulnerabilities in the firmware code itself which allow an attacker to
+execute arbitrary code with the capabilities of the firmware and (2)
+firmware with intentional Trojan (backdoor) capabilities.
+
+%Vulnerabilities in the firmware code allow an attacker to execute
+%arbitrary code with the permissions of the firmware. The class of
+%vulnerabilities are not unique to firmware, as they are the same class
+%of vulnerabilities that infect traditional binary software (buffer
+%overflows, logic flaws, etc.). However, firmware typically has full
+%control over the entire computing system, therefore vulnerabilities in
+%firmware are much more severe than those in the application layer. In
+%addition, if an attacker is able to persist their exploit --- because the
+%compromise occurs at such a low layer of the computing system --- the
+%attacker can persist \emph{even after reinstalling the operating
+%	system or application}.
+
+%Faris:
+Vulnerabilities in the firmware code allow an attacker to execute
+arbitrary code with the permissions of the firmware. The class of
+vulnerabilities are not unique to firmware, as they are the same class
+of vulnerabilities that infect traditional binary software (buffer
+overflows, logic flaws, etc.). However, firmware typically has full
+control over the entire computing system, therefore vulnerabilities in
+firmware are much more severe than those in the application layer. In
+addition, the attacker may remain in the system because the
+compromise occurs at such a low layer of the computing system so the
+exploit persists \emph{even after reinstalling the operating
+	system or application}.
+
+%The attack surface for firmware varies depending on where it is used.
+%In the most general case, the attacks can come from the upper layers,
+%specifically the Operating System or Hypervisor layers, as these are
+%the layers of the computing stack that can directly communicate with the
+%firmware. However, other firmware can be attacked either directly
+%through the Hardware layer itself, or through the data being
+%transmitted through the firmware from the hardware. For instance, a
+%network card firmware could be attacked through network traffic that
+%the firmware processes.
+
+%Faris:
+The attack surface for firmware varies depending on where it is used.
+In the general case, the attacks can come from the upper layers,
+specifically the Operating System or Hypervisor layers, as these are
+the layers of the computing stack that can directly communicate with the
+firmware. However, other firmware can be attacked either directly
+through the Hardware layer, or through the data being
+transmitted through the firmware from the hardware. For instance, a
+network card firmware could be attacked through network traffic that
+the firmware processes.
+
+Traditional vulnerability analysis is not effective on firmware, due
+to the specificity of the firmware for the specific hardware and the
+lack of a well-defined operating system abstraction for the firmware
+code. In essence, each firmware is a custom binary with little in the
+way of assumptions that vulnerability analysis can take advantage of.
+
+%Trojan capabilities in the firmware itself can allow a system to be
+%remotely controlled or accessed. This is similar to a vulnerability,
+%in that an attacker is able to control the firmware, however the key
+%difference lies in the fact that Trojan behavior will be intentionally
+%obfuscated by an attacker, whereas a vulnerability is accidentally
+%introduced by a benign developer. Therefore, Trojan behavior is 
+%more difficult to detect, because there is an active adversary who is
+%intentionally obfuscating the behavior.
+
+%Faris:
+Trojan capabilities in the firmware can allow a system to be
+remotely controlled or accessed. This is similar to a vulnerability,
+in that an attacker is able to control the firmware, however the key
+difference lies in the fact that Trojan behavior can be intentionally
+obfuscated by an attacker, whereas a vulnerability can accidentally
+introduced by a benign developer. Therefore, Trojan behavior is 
+difficult to detect.
+
+%Vulnerabilities in the firmware and Trojan capabilities are related
+%because both require \emph{understanding} the behavior and semantics
+%of the firmware code. This is why reverse engineering is such a
+%critical component of securing the threats against the firmware layer:
+%in order to detect either vulnerabilities or Trojan functionality, an
+%automated tool must first understand the semantics of the particular
+%firmware/hardware combination.
+
+%Faris:
+Vulnerabilities in the firmware and Trojan capabilities are related
+because both require \emph{understanding} the behavior and semantics
+of the firmware code. This is why reverse engineering is such a
+critical component of securing the firmware layer against the threats.
+In order to detect vulnerabilities or trojan functionality, an
+automated tool must understand the semantics of the particular
+firmware/hardware combination.
+
+\subsection{Research}
+%Faris: Some fixes below. 
+Researchers have attempted to identify real-world attacks on firmware
+code~\cite{tsow2006, cui2010, duflot2011, basnight2013, cui2013,
+	zaddach2013, zaddach2013a, maskiewicz2014, stuttgen2015}. These
+attacks are particularly effective, and researchers have shown that
+these are able to persist through a reboot and cannot be detected by
+the upper layers.
+
+Other approaches are about how to prevent the tampering of
+firmware~\cite{adelstein2002, zhou2007, zhou2009}. These protect from
+the threat of tampering or altering the firmware, however it does not
+address vulnerable firmware.
+
+Another approach includes runtime monitoring into the
+firmware, so that compromises can be detected~\cite{cui2010a,
+	cui2011}. This approach is further complicated by the lack of
+abstractions in the firmware environment.
+
+Recent work has looked into the problem of updating the design of the
+firmware so that it has security features, such as memory handling and
+avoiding information leakage~\cite{koeberl2014, kauer2007}. However,
+forcing firmware developers to use such features is an open issue.
+
+Very few work has looked into how to reverse engineer and understand the
+firmware~\cite{zaddach2014}. This problem is of particular interest,
+as understanding the semantics of the firmware and how it interacts
+with hardware is a key challenge.
+
+Recent work has looked into how to detect vulnerabilities in
+firmware~\cite{davidson2013, costin2014, shoshitaishvili2015}. One of
+these has an excellent website\footnote{\url{http://firmware.re}}
+that contains the results of their reverse engineering and
+vulnerability analysis~\cite{costin2014}.
+
+\subsection{Suggested Focus Areas}
+%Faris: Some fixes below.
+Of all the layers, firmware has received the least attention from the
+research community. Therefore, we believe that focus on this research
+area will yield significant dividends.
+
+The major challenge with preventing firmware vulnerabilities is how to decouple
+firmware from the hardware, potentially using emulation of the
+hardware, in order to aid the analysis of the firmware.
+
+In a similar vein, there is need for automated reverse engineering tools,
+including analysis of memory layouts, code modules, even answering the
+simple question of what is code in a firmware binary and what is data.
+
+Finally, even if we find vulnerabilities or Trojan logic in a
+firmware, there is still the unanswered question of how to patch or
+update the functionality of the firmware while still guaranteeing the
+behavior of the firmware. This is particularly complicated due to the
+real-time requirements of the firmware and the potentially concurrent
+executions.
 
@@
 
-% Start of "Sample References" section
-
-\section{Typical references in new ACM Reference Format}
-A paginated journal article \cite{Abril07}, an enumerated
-journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
-a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
-\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
-followed by the same example, however we only output the series if the volume number is given
-\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
-a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
-in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
-an article in a proceedings (of a conference, symposium, workshop for example)
-(paginated proceedings article) \cite{Andler79}, a proceedings article
-with all possible elements \cite{Smith10}, an example of an enumerated
-proceedings article \cite{VanGundy07},
-an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
-a master's thesis: \cite{anisi03}, an online document / world wide web resource \cite{Thornburg01}, \cite{Ablamowicz07},
-\cite{Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
-and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
-work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
-\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
-'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
-Boris / Barbara Beeton: multi-volume works as books
-\cite{MR781536} and \cite{MR781537}.
-
 % Appendix
@@
 \bibliographystyle{ACM-Reference-Format-Journals}
-\bibliography{acmsmall-sample-bibfile}
 \bibliography{papers}

### 2016-08-21 paper.tex
think i fixed the biblio issue

@@
 analysis to identify the malicious Trojan behavior~\cite{lee2005,
-agrawal2007, banga2008, banga2008a, jin2008, du2010, narasimhan2011,
-narasimhan2012, cao2013, davoodi2013, dworak2013, forte2013,
-liu2013, liu2014, bao2015}. This research uses power analysis, heat
+agrawal2007,banga2008,banga2008a,jin2008,du2010,narasimhan2011,
+narasimhan2012,cao2013,davoodi2013,dworak2013,forte2013,
+liu2013,liu2014,bao2015}. This research uses power analysis, heat
 analysis, or other side-effects to determine that the hardware is
@@
 Other research has demonstrated the attacks that are possible by
-actually implementing Trojan behavior~\cite{king2008, jin2010,
-sturton2011, liu2013, jin2014, zhang2014}. This line of research
+actually implementing Trojan behavior~\cite{king2008,jin2010,
+sturton2011,liu2013,jin2014,zhang2014}. This line of research
 shows that Trojan logic is feasible for an attack.
@@
 effectiveness of the side-channel analysis~\cite{chakraborty2008,
-li2008, salmani2009, hicks2010, rosenfeld2011, waksman2011,
-bhunia2013, davoodi2013, rajendran2013, rolt2014}. This line of
+li2008,salmani2009,hicks2010,rosenfeld2011,waksman2011,
+bhunia2013,davoodi2013,rajendran2013,rolt2014}. This line of
 research attempts to rethink designing the hardware itself. However, a
@@
 has investigated the problem of securing JTAG access to
-hardware~\cite{rosenfeld2010, ren2015}. The main idea is to allow only
+hardware~\cite{rosenfeld2010,ren2015}. The main idea is to allow only
 authenticated users to have access to the JTAG functionality, however
@@
 which provide a good overview of the field~\cite{wang2008,
-tehranipoor2010, guin2014, guin2014a, guin2014b}.
+tehranipoor2010,guin2014,guin2014a,guin2014b}.
 
@@
 Researchers have attempted to identify real-world attacks on firmware
-code~\cite{tsow2006, cui2010, duflot2011, basnight2013, cui2013,
-	zaddach2013, zaddach2013a, maskiewicz2014, stuttgen2015}. These
+code~\cite{tsow2006,cui2010,duflot2011,basnight2013,cui2013,
+zaddach2013,zaddach2013a,maskiewicz2014,stuttgen2015}. These
 attacks are particularly effective, and researchers have shown that
@@
 Other approaches are about how to prevent the tampering of
-firmware~\cite{adelstein2002, zhou2007, zhou2009}. These protect from
+firmware~\cite{adelstein2002,zhou2007,zhou2009}. These protect from
 the threat of tampering or altering the firmware, however it does not
@@
 firmware so that it has security features, such as memory handling and
-avoiding information leakage~\cite{koeberl2014, kauer2007}. However,
+avoiding information leakage~\cite{koeberl2014,kauer2007}. However,
 forcing firmware developers to use such features is an open issue.
@@
 Recent work has looked into how to detect vulnerabilities in
-firmware~\cite{davidson2013, costin2014, shoshitaishvili2015}. One of
+firmware~\cite{davidson2013,costin2014,shoshitaishvili2015}. One of
 these has an excellent website\footnote{\url{http://firmware.re}}

### 2016-08-21 paper.tex
definitely solved the biblio prob

@@
 Most research into hardware Trojans deal with using side-channel
-analysis to identify the malicious Trojan behavior~\cite{lee2005,
-agrawal2007,banga2008,banga2008a,jin2008,du2010,narasimhan2011,
-narasimhan2012,cao2013,davoodi2013,dworak2013,forte2013,
-liu2013,liu2014,bao2015}. This research uses power analysis, heat
+analysis to identify the malicious Trojan behavior~\cite{lee2005,agrawal2007,banga2008,banga2008a,jin2008,du2010,narasimhan2011,narasimhan2012,cao2013,davoodi2013,dworak2013,forte2013,liu2013,liu2014,bao2015}.
+This research uses power analysis, heat
 analysis, or other side-effects to determine that the hardware is
@@
 Other research has demonstrated the attacks that are possible by
-actually implementing Trojan behavior~\cite{king2008,jin2010,
-sturton2011,liu2013,jin2014,zhang2014}. This line of research
+actually implementing Trojan behavior~\cite{king2008,jin2010,sturton2011,liu2013,jin2014,zhang2014}. This line of research
 shows that Trojan logic is feasible for an attack.
@@
 in order to improve the testability of the chip or to increase the
-effectiveness of the side-channel analysis~\cite{chakraborty2008,
-li2008,salmani2009,hicks2010,rosenfeld2011,waksman2011,
-bhunia2013,davoodi2013,rajendran2013,rolt2014}. This line of
+effectiveness of the side-channel analysis~\cite{chakraborty2008,li2008,salmani2009,hicks2010,rosenfeld2011,waksman2011,bhunia2013,davoodi2013,rajendran2013,rolt2014}. This line of
 research attempts to rethink designing the hardware itself. However, a
@@
 Finally, there are several surveys conducted on hardware security,
-which provide a good overview of the field~\cite{wang2008,
-tehranipoor2010,guin2014,guin2014a,guin2014b}.
+which provide a good overview of the field~\cite{wang2008,tehranipoor2010,guin2014,guin2014a,guin2014b}.
 
@@
 Researchers have attempted to identify real-world attacks on firmware
-code~\cite{tsow2006,cui2010,duflot2011,basnight2013,cui2013,
-zaddach2013,zaddach2013a,maskiewicz2014,stuttgen2015}. These
+code~\cite{tsow2006,cui2010,duflot2011,basnight2013,cui2013,zaddach2013,zaddach2013a,maskiewicz2014,stuttgen2015}. These
 attacks are particularly effective, and researchers have shown that
@@
 Another approach includes runtime monitoring into the
-firmware, so that compromises can be detected~\cite{cui2010a,
-cui2011}. This approach is further complicated by the lack of
+firmware, so that compromises can be detected~\cite{cui2010a,cui2011}.
+This approach is further complicated by the lack of
 abstractions in the firmware environment.

### 2016-08-21 paper.tex
added till hypervisors, fixed biblio till hypervisors

@@
 
-\section{Performance Evaluation}
+\section{Bus and Interconnect}
 
-During all the experiments, the Geographic Forwarding (GF)
-[Akyildiz 2001] routing protocol is used. GF exploits geographic
-information of nodes and conducts local data-forwarding to achieve
-end-to-end routing. Our simulation is
-configured according to the settings in
-Table~\ref{tab:one}. Each run lasts for 2 minutes and
-repeated 100 times. For each data value we present in the results,
-we also give its 90\% confidence interval.
-% Table
-\begin{table}%
-\tbl{Simulation Configuration\label{tab:one}}{%
-\begin{tabular}{|l|l|}
-\hline
-TERRAIN{$^a$}   & (200m$\times$200m) Square\\\hline
-Node Number     & 289\\\hline
-Node Placement  & Uniform\\\hline
-Application     & Many-to-Many/Gossip CBR Streams\\\hline
-Payload Size    & 32 bytes\\\hline
-Routing Layer   & GF\\\hline
-MAC Layer       & CSMA/MMSN\\\hline
-Radio Layer     & RADIO-ACCNOISE\\\hline
-Radio Bandwidth & 250Kbps\\\hline
-Radio Range     & 20m--45m\\\hline
-\end{tabular}}
-\begin{tabnote}%
-\Note{Source:}{This is a table
-sourcenote. This is a table sourcenote. This is a table
-sourcenote.}
-\vskip2pt
-\Note{Note:}{This is a table footnote.}
-\tabnoteentry{$^a$}{This is a table footnote. This is a
-table footnote. This is a table footnote.}
-\end{tabnote}%
-\end{table}%
+The Bus and Interconnect is an abstract layer of the computing system
+which allows the firmware and hardware to communicate with each other.
+Different computing architectures have different technical
+implementations of the bus and interconnect. For instance, the
+communication network in the automotive system, which has received
+significant attention from the research community, uses a CAN bus to
+communicate. Modern motherboards use either a Northbound or Southbound
+bus --- such as USB --- to communicate between the CPU and other peripheral devices.
+
+\subsection{Threats}
+
+Depending on the technical architecture of the bus and interconnect
+layer, the bus can be used either purely for communication or for
+other functionality as well. For instance, the CAN bus is only used to
+allow different hardware components to communicate. However, on modern
+motherboards, techniques such as Direct Memory Access (DMA) allow a
+peripheral device that is connected to the bus to directly access the
+system's memory, without going through the CPU.
+
+In the case of the bus and interconnect being used as a communication
+medium, then the classic threats are the same security concerns as
+other mediums of communication: communication integrity and
+confidentiality. A malicious device on the bus can impersonate another
+device and send messages on the bus as the impersonated device. On an
+automotive system, researchers have demonstrated that by compromising
+a single device attached to the system's CAN bus allows an attacker to
+have complete control over all functions of the automobile, by
+impersonating messages on the CAN bus. In addition to impersonating
+messages, an attacker on the bus and interconnect layer can listen to
+messages on the bus, potentially eavesdropping on sensitive messages.
+
+If the bus and interconnect layer offers advanced functionality, such
+as the DMA example, then the devices that an attacker is able to connect
+to the bus could take advantage of this functionality. In the case of
+DMA, this means that any device on the bus has full access to the
+physical memory of the system, therefore the attacker has complete
+control over the upper layers of the computing system (hypervisor,
+operating system, and application).
+
+The difficulty of addressing the threats of the bus and interconnect
+layer is the diversity and specificity of the specific bus and
+interconnect technology. Often, these were created without
+consideration for security needs, therefore, as a communication
+mechanism they lack authentication from the design. In addition, as
+the bus and interconnect layer cause different hardware to
+communicate, all hardware must agree on the protocol to talk on the
+bus. Therefore, adding security to the bus and interconnect protocol
+is difficult in legacy systems.
+
+\subsection{Research}
+
+A major area of research in the Bus and Interconnect layer is work
+that shows the dangers of a bus and interconnect that allows malicious
+devices to either impersonate devices on the bus or leverage bus
+capabilities, particularly in the automotive area~\cite{bonkoski2013,checkoway2011,hoppe2008,koscher2010,rouf2010,sang2010}.
+
+The other major area of research is to apply encryption to devices on
+the bus and interconnect~\cite{groza2012,herrewege2011,jiang2012,schweppe2011,stewin2014,szilagyi2010}.
+In this way, malicious devices on the bus cannot impersonate other devices. 
+However, a key problem with this research direction is how to distribute the keys of
+the devices, which is typically done by distributing them in hardware.
+Unfortunately, this approach is brittle, as it disallows new devices
+from accessing the bus and interconnect.
+
+Other research has investigated how to ensure the integrity of devices
+attached to the bus and interconnect using attestation
+techniques~\cite{li2010,li2011}. These techniques attempt to ensure
+that only authorized and unmodified devices are attached to the bus
+and interconnect. However, this research has the problem of how to
+allow third-party devices to access the bus.
+
+A key problem is how to perform secure updates to devices attached to
+the bus and the interconnect~\cite{larson2008}. This is a challenge
+that also has similarities with the firmware layer.
+
+In order to discover malicious communication, researchers have
+utilized data flow tracking techniques to discover how the data flows
+between devices connect to the bus and interconnect
+layers~\cite{schweppe2012}. Using these techniques, the researchers
+are able to identify malicious flows or potential vulnerabilities by
+analyzing the way that data flows throughout the system.
+
+Other researchers have explored modeling either the system itself
+or the behavior of the devices on the bus~\cite{stewin2013a,drolia2011}. 
+If the device behavior or the system behavior does not
+match the learned model, then an alert is triggered. In this way,
+these techniques are similar to anomaly detection techniques, with
+similar benefits and drawbacks.
+
+In order to address the issue of devices on the bus able to access
+capabilities unauthorized, researchers have tackled the similar
+problem of reducing the Trusted Computing Base, essentially reducing
+the functionality of the bus and interconnect
+layer~\cite{vasudevan2012,zhang2013,zhou2009}. In this way, if the
+functionality is not there on the bus, then a malicious device cannot
+take advantage of it.
+
+Another approach is to leverage hardware features of the underlying
+bus and interconnect system to reduce the impact of an
+attack~\cite{wolf2012,wojtczuk2011}. However, as these hardware
+functionalities must be implemented, it's impact in legacy systems is
+limited.
+
+Finally, other research in this area has performed surveys of the
+research on this area~\cite{kleberger2011,studnia2013,thom2008,wolf2004,wolf2007,wright2011,zhao2002}.
+
+\subsection{Suggested Focus Areas}
+
+Further research in this area should focus on enabling better
+monitoring of the runtime of devices attached to the bus or
+interconnect. There is also the need to support logging of the bus or
+interconnect activity. This will enable detection of a malicious
+device on the bus or interconnect.
+
+Another promising area of research is the automated filtering of
+the bus communication, potentially allowing an administrator to define
+a policy for a specific bus or interconnect system. This would be
+particularly effective if it can be implemented on legacy bus and
+interconnect systems, with no modification to the devices.
+
+A different approach would be to secure the bus by only allowing
+authorized devices to attach to the bus. This could be a white-listing
+based approach or other policy-driven approach. However, the key
+question is how to identify a device, particularly a third-party
+device.
+
+While there has been some research on encrypting the communication on
+the bus, most has focused on pre-distributing the keys to devices on
+the bus. This approach is brittle, as new devices cannot be added to
+the bus easily. Therefore, further approaches should be investigated,
+such as using attribute-based encryption to be able to encrypt and
+distribute messages on the bus without prior sharing of keys.
+
+\section{Hypervisor}
+
+The purpose of the hypervisor layer is to provide virtualization
+techniques for the upper layers of the computing stack. In essence,
+the Hypervisor layer provides a virtualization or emulation of the
+lower layers of the computing stack, so that multiple Operating
+Systems, which each assume that they are the sole users of the
+underlying hardware, can run concurrently on the same hardware.
+
+For discussing the Hypervisor layer, we use the term \emph{host} to
+define the hypervisor itself (which, in some cases can be an Operating
+System) and the \emph{guest} is the Operating System that is running
+on the hypervisor. A host can run multiple guest Operating Systems.
+
+Hypervisors have a number of benefits, from allowing greater sharing
+and usage of hardware resources, to improving the isolation of the
+Operating Systems. The security benefits of using Hypervisors comes
+from each guest Operating System being isolated from the other. In
+theory, this means that a compromise of a guest Operating System will
+be isolated to only that Operating System, and the attack should not
+spread to other guest Operating Systems.
+
+\subsection{Threats}
+
+The central threat to the Hypervisor layer is from attackers in a
+guest Operating System exploiting a vulnerability in the hypervisor
+itself to gain access to the actual hardware or to propagate their
+attack to the other guest Operating Systems that are running in the
+hypervisor. As the Hypervisor layer is essentially emulating the
+hardware for the guest operating systems, it has a large attack
+surface. Furthermore, modern hypervisors, in an attempt to increase
+the performance of the hypervisor, frequently allow guest Operating
+Systems direct access to the hardware. If the hardware does not
+provide for full isolation (i.e., it is not hypervisor-aware), then an
+attacker can leverage this hardware to tamper or extract sensitive
+information from other guest Operating Systems.
+
+In modern hypervisors, the hypervisors frequently provide isolation
+from a well-behaved guest Operating System, as the guest Operating
+System is assumed to be benign. This assumption can be unfounded, as
+the guest Operating System can be hostile and can attempt to exploit
+or crash the hypervisor.
+
+Another threat actually stems from the Hypervisor itself being
+untrusted. In the cloud computing environment, the guest Operating
+System does not trust the underlying Hypervisor, as it is controlled
+by a third-party. In this case, the Hypervisor can spy on the guest
+or alter the memory of the guest.
+
+\subsection{Research}
+
+One area of research is to leverage the vantage point of the
+Hypervisor to monitor the guest for security
+violations~\cite{karger1991,barham2003}, intrusion
+detections~\cite{dunlap2002,garfinkel2003a,kourai2005,chen2008,garfinkel2003b},
+ checking the integrity of the guest Operating
+system~\cite{seshadri2006,mccune2010}, and identifying rootkits present
+in the guest Operating System~\cite{jones2008,grace2010}. This work
+assumes that the guest can be compromised and that the Hypervisor is
+trusted, which, due to vulnerabilities in the Hypervisor, is not
+always true.
+
+A significant problem in monitoring or otherwise inspecting a guest
+Operating System from the hypervisor layer is the \emph{semantic gap}
+problem, where the Hypervisor has no outside knowledge of the
+guest-specific Operating System constructs. Therefore, research has
+investigated how to overcome this semantic gap~\cite{jones2006,christodorescu2009,srinivasan2011}.
+
+Further work in this line has approached the problem of how to monitor
+the hypervisor itself by leveraging hardware features~\cite{azab2010}.
+
+In order to build a secure Hypervisor, the lower layers of the
+computing system must also be trusted. Therefore, research into
+establishing a Trusted Computing Base, which validates the integrity
+of the lower levels for the guest Operating System~\cite{arbaugh1997,sailer2004,singaravelu2006}
+ is important for ensuring the security
+of the Hypervisor.
+
+One problem of using the Hypervisor as a security primitive is that it
+is part of the Trusted Computing Base for the upper layers of the
+computing platform. Therefore, attempts have been made to reduce the
+trusted base of the Hypervisor itself~\cite{murray2008,shinagawa2009,colp2011,wang2012}.
+
+Other research has looked at creating a secure
+Hypervisor~\cite{robin2000,belay2012,szefer2011,vasudevan2013} with
+the possibility of adding functionality to the Hypervisor to support
+closed guest Operating Systems~\cite{garfinkel2003}, offering
+applications running in the guest Operating System to hide information
+from the guest Operating System~\cite{chen2008,yang2008,jin2011,szefer2012,azab2011,zhang2011,jin2011a,szefer2011a},
+ secure information flow between the guest Operating System drivers and I/O
+devices~\cite{cheng2013}, or designing the hypervisor so that it can
+prove that it did not tamper with the guest Operating
+System~\cite{gu2011}.
+
+Another approach to improve the security of the Hypervisor is to
+remove the duplicated components from the guest Operating System
+layer~\cite{kivity2014,madhavapeddy2015}, which effectively reduces
+the attack surface of the Operating System layer.
+
+Another interesting approach is to run traditional network middleware
+boxes, such as firewalls and Intrusion Detection Systems, as a guest
+Operating System of a Hypervisor~\cite{martins2014}.
+
+Finally, there have been attempts to introduce traditional
+application-level security features to Hypervisors, in order to reduce
+or prevent vulnerabilities, such as control-flow
+integrity~\cite{wang2010} or manual vulnerability
+analysis~\cite{Perez-Botero2013}.
+
+\subsection{Focus Areas}
+
+As the hypervisor is typically a monolithic trusted entity, research
+should be performed to reduce the trust in the hypervisor. One avenue
+could be to have fine grained access control within the hypervisor, so
+that different policies can be set to the guests hosted by the
+hypervisor.
+
+In a similar area, we believe that further research into reducing the
+attack surface of the hypervisor is necessary. Specifically, the
+hypervisor can be customized based on the applications and workload
+necessary.
+
+Hypervisors, despite their obvious benefits, have not been applied to
+real-time systems. Further research must be done to allow reliable and
+robust hypervisors in a constrained real-time computing environment.
+
+Testing hypervisors is currently a difficult and essentially manual
+process. There is currently no equivalent to fuzz testing a hypervisor
+implementation, possibly because there are no models for attacks
+against a hypervisor. Further research in increasing the testability
+and automated analysis of hypervisor reliability would improve security.
 

### 2016-08-21 paper.tex
added till os, fixed biblio till os

@@
 
+\section{Operating Systems}
+
+The Operating System layer is responsible for ensuring that multiple
+applications are able to concurrently run, without interfering with
+each other. In addition, the Operating System layer is also
+responsible for providing a unified abstraction of the underlying
+hardware layer for the application layer. Therefore, in terms of
+security, the Operating System layer has more privileges than the
+application layer, as the Operating System controls the execution and
+scheduling of processes, the access control policies that allow one
+application to hide information/data from the other processes.
+
+\subsection{Threats}
+
+The main threat to the Operating System layer is an attacker
+compromising the security of the Operating System. There are two main
+avenues to how this compromise can occur: (1) an attacker exploits a
+vulnerability in the Operating System from the Application layer or
+(2) an attacker compromises a lower layer (for instance, modifying the
+Operating System code directly to the hardware's Hard Drive).
+
+Vulnerabilities in the Operating System allow an attacker to have
+complete control over the other application running in the operating
+system. In addition, the attacker can use their vantage point at the
+operating system to attempt to escalate their attack to the hypervisor
+or lower layers, such as the firmware. Vulnerabilities in the
+Operating System result from programming errors that allow an attacker
+to gain control of the Operating System code or data. These
+vulnerabilities are traditional memory corruption vulnerabilities or
+even logic flaws.
+
+A main concern in Operating System security is the presence of
+rootkits, which are malicious modifications to the Operating System
+which an attacker uses to achieve persistence on the system. In
+addition, attackers use rootkits to hide from other applications on
+the system. For instance, the rootkit will modify the operating system
+code such that the attacker's processes and files are hidden from the
+administrator or defensive software installed on the Operating System,
+such as an anti-virus solution.
+
+Direct access to the hardware can allow an attacker to compromise the
+Operating System. It is in this way that an attacker can propagate
+their compromise from the lower layers of the computing platform to
+the upper layers. The attacker has similar capabilities as when
+exploiting a vulnerability in the Operating system.
+
+\subsection{Research}
+
+Research into Operating System security has a long and rich history,
+which has looked at design flaws in Operating
+Systems~\cite{attanasio1973}. Further research looked at reducing the
+permissions of the \emph{root} users, essentially compartmentalizing
+the privilege of the Operating System~\cite{kamp2000,kurmus2014a}.
+
+Rootkits attempt to persist the attacker's compromise of the operating
+system and allow them to hide their malicious behavior from other
+applications on the system. Therefore, rootkit detection is an
+important research topic in Operating System
+security~\cite{petroni2004,wang2005,petroni2007,payne2008,sharif2009,yin2010,bianchi2012}.
+Another approach is to prevent
+the rootkit from altering the Operating System
+kernel~\cite{riley2008,wang2009,grace2010,gadaleta2011}.
+
+Vulnerability analysis techniques have been used to discover
+vulnerabilities in the Operating System kernel before a hacker is able
+to use and exploit those vulnerabilities. Static analysis techniques
+have been used~\cite{chou2001,chen2011}.
+
+Other vulnerability defenses include defenses for specific
+vulnerability classes, such as buffer overflows~\cite{dalton2008}.
+Randomization techniques attempt to alter the environment of the
+operating system in such a way that the system continues to function
+but the attacker's exploit fails~\cite{lin2009,giuffrida2012}.
+Control-flow integrity techniques have been applied to Operating
+System code as well~\cite{li2011a}.
+
+Other research into Operating System security has looked at how to
+protect applications from a malicious or compromised Operating
+System~\cite{ta-min2006,onarlioglu2013,li2014}.
+
+% either driver code execute or change `and' to `is'
+As driver code executes in kernel space, and is a frequent source of
+vulnerabilities~\cite{kadav2012}, researchers have looked into how to
+make drivers more robust~\cite{ganapathy2008,hruby2012} or isolated
+from the rest of the kernel~\cite{boyd-wickizer2010}.
+
+Attack papers show that it is possible to compromise the kernel
+without injecting any additional code, using Return Oriented
+Programming techniques~\cite{hund2009}.
+
+A completely different approach is to formally prove the correctness
+of the Operating System kernel~\cite{klein2009}.
+
+Updating Operating Systems with security updates is frequently
+difficult, as the Operating System must be restarted. However,
+researchers have explored how to apply an update to the Operating
+System without rebooting the system~\cite{arnold2009}.
+
+Some work has been done on removing the operating system
+entirely~\cite{madhavapeddy2010} or trimming/tailoring parts of the
+Operating System~\cite{kurmus2011,madhavapeddy2013,howell2013,kurmus2013,kurmus2014}.
+This has the security benefit of reducing
+the attack surface of the Operating System.
+
+\subsection{Suggested Focus Areas}
+
+We believe that, while significant research has been done on the
+Operating System layer, there still remains work to be done in this
+area. In particular, as the Operating System is frequently in contact
+with malicious applications, it must be hardened from attack.
+
+Therefore, we believe that further research into the line of trimming
+or tailoring the operating system is important, such that the attack
+surface is significantly reduced.
+
+We also believe that non-control data-flow attacks will be the new
+wave of attacks against Operating Systems. Therefore, we need new
+analysis techniques, both static and dynamic, to discover these
+vulnerabilities that do not require altering the control flow of the
+Operating System.
+
+A critical component of the Operating System is to handle the
+execution of concurrent applications. Therefore, attacks that take
+advantage of the concurrency in the Operating System are unexplored.
+Research must be done to create new models for these attacks.
+
 \section{Conclusions}

### 2016-08-22 paper.tex
fixed biblio till app layer

@@
 % Page heads
-\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}
+\markboth{ et al.}{Total Platform Cyber Protection}
 
@@
 
-\acmformat{Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic,
-and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially
-designed for  wireless sensor network applications.}
+\acmformat{, 2016. Total Platform Cyber Protection.}
 % At a minimum you need to supply the author names, year and a title.
@@
 Research must be done to create new models for these attacks.
+
+\section{Application Layer}
+
+The application layer is the layer of the computing stack that
+performs that actual useful computation or offers the computing
+service. It is essentially at the top of the computing stack, as the
+lower layers of the computing stack provide abstractions so that
+applications are easier to write and the upper layer (networking) is
+used by the application layer to communicate to other applications.
+
+At the application layer, we consider all the following types of
+``applications'' under the application layer umbrella: GUI
+applications, command-line applications, server batch process
+applications, web applications, or mobile applications. Also, all
+programming languages used to develop the applications are also in
+scope, from compiled languages such as C or C++ to scripting languages
+such as Python and Ruby. In short, all types of computing applications
+are included in the application layer.
+
+Due to abstractions provided by the lower layers, applications
+typically have lesser privilege than the lower layers of the computing
+stack. This is due to the abstractions provided to the application
+layer, for instance an application cannot access the underlying hard
+drive directly, but must use the Operating System's provided
+interface.
+
+Applications developed are typically very complex to understand
+completely. This is due in part to the reliance on the abstractions
+provided by the lower computing layers (as these abstractions must be
+fully understood to understand how the application uses them). Another
+aspect is that applications frequently use shared libraries or other
+middleware, which are also considered to be part of the application
+layer. In this way, multiple different applications can all use the
+same underlying library. As the application's behavior depends on the
+behavior of \emph{all} used libraries and abstractions, this means
+that understanding even a simple application is dependent on the
+complexity of the used abstractions.
+
+\subsection{Threats}
+
+Applications, by definition, must be accessible by users in
+order to perform their useful computation. Due to this accessibility,
+the application layer, of all the other computing layers, is the entry
+point for attacks or exploits.
+
+At a high level the threat to the application layer is attackers being
+able to influence the system to perform computation of their choosing.
+There are two main vectors for an attacker to influence the system:
+(1) software vulnerabilities in an application that can allow an
+attacker to exploit the vulnerability to either perform arbitrary
+computation or otherwise influence the application and (2) a malicious
+application (typically described as a Trojan application) that the
+attacker tricks the user into installing and running on the system,
+usually by using social engineering techniques.
+
+Vulnerabilities are flaws or bugs in the application (or any software
+for that matter) that allows an attacker to compromise the security of
+the application. The key difference from normal flaws or bugs in
+software is that vulnerabilities allow the program to deviate from the expected behavior. These vulnerabilities range the gamut of possible causes
+such as programming errors or configuration mistakes. They can be in
+the actual, custom application software or in any of the libraries
+that the application uses. Some vulnerabilities are specific to a
+given programming language, while others are more broad and can any
+affect application. Examples of common vulnerabilities are: buffer
+overflow, heap overflow, format string, off-by-one, integer overflow,
+time-of-check-time-of-use, cross-site scripting, SQL injection, and
+logic flaws.
+
+Trojan applications attempt to control or manipulate the system by
+tricking a user into executing them. They frequently do this by using
+social engineering techniques to entice the user to execute them. The
+application may appear innocuous, such as a flashlight application on
+a mobile device, but may in fact include malicious functionality that
+the user does not want.
+
+Using either of these two techniques, the attackers goal is to move
+from an external position to having influence on the executing system.
+The best scenario for the attacker is to completely control the
+execution of the application. From here, the attacker can now
+arbitrarily access any resource that the application can access. For
+instance, this could mean reading sensitive data, changing sensitive
+data, deleting sensitive data, or using the vantage point of the
+application to further the compromise to the lower levels of the
+computing platform or even to other applications on the network.
+
+Applications have a large and complex attack surface, which increases
+the chances of a developer accidentally introducing a vulnerability.
+The attack surface of the application is increased by the transitive
+closure of the libraries that the application uses (and that those
+libraries use). A vulnerability in a library is a vulnerability in the
+application that uses that library. Therefore, the complexity and
+increased attack surface of applications are a threat, as they allow
+increases in vulnerabilities.
+
+\subsection{Research}
+
+Security of the application layer has received significant attention
+from the research community. Because of this, we have divided the
+research into the following categories: attack surface reduction, prevent control-flow
+hijacking, logic and implementation flaws, mitigation and recovery,
+modeling and monitoring execution, and protection against malicious
+code.
+
+\subsubsection{Attack surface reduction}
+
+The idea behind this subarea of application layer security research is
+to reduce the attack surface of the application. This line of research
+views vulnerabilities as a probability, and by reducing the attack
+surface, the odds of an attack being able to exploit a
+vulnerability successfully is reduced as well.
+
+Attempts have been made to measure an application's attack
+surface~\cite{howard2003,manadhata2007,manadhata2008,ruprecht2014}.
+This has the benefit of attempting to quantify the attack surface of
+an application, which in itself is a difficult problem.
+
+Several research works attempt to reduce the attack surface. Some
+approaches try to statically analyze for deadcode, then restrict the
+execution of the deadcode during runtime~\cite{kurmus2011}. Other ways
+attempt to leverage compile-time configurability to reduce the attack
+surface~\cite{tartler2012,kurmus2013,ruprecht2014}. A different
+approach is to reduce the attack surface by partitioning the
+application into different privilege levels~\cite{barth2010,xu2012}.
+Other approaches attempt to restrict the communication between
+applications, in order to reduce opportunities for the attacker to
+input data to the system~\cite{kantola2012}.
+
+Finally, other approaches attempt to quantify the attack surface and
+measure the actual reduction~\cite{kurmus2014}.
+
+\subsubsection{Prevent control-flow hijacking}
+
+Modern attackers, exploiting binary applications such as browsers,
+media players, or PDF readers, attempt to hijack the control-flow of
+the application. If an attacker successfully hijacks the control-flow
+of the application, then she has complete control over the execution
+of the application. Many vulnerability classes enable control-flow
+hijacking, specifically memory corruption vulnerabilities such as
+buffer overflows, heap overflows, and format strings~\cite{szekeres2013}.
+
+The goal of this line of research is to prevent the attacker from
+successfully hijacking the control-flow of the application. In this
+way, this line of research can be seen as a mitigation technique.
+
+Significant progress in this area started with the idea of
+control-flow integrity, essentially ensuring dynamically at runtime
+that the control-flow of the application does not
+deviate~\cite{abadi2005,abadi2009}. The first implementations were
+deemed infeasible, so later research attempted to improve the
+efficiency of control-flow integrity~\cite{zhang2013,tice2014,bletsch2011,zhang2013,zhang2013a}.
+
+Researchers have advanced the state of attacks against control-flow
+hijacking. The most advanced of these is return-oriented programming
+(ROP), in which an attacker reuses bits of code from the original
+application, so that the attacker can hijack the control-flow of the
+application without introducing new code~\cite{roemer2012,bletsch2010,tran2011}. 
+Researchers have also improved these ROP
+techniques~\cite{checkoway2010,buchanan2008,kayaalp2012,chen2012,goktas2014,schwartz2011}
+ and applied them to a different
+architecture~\cite{davi2012}.
+
+Other research has looked at defending ROP attacks~\cite{pappas2012,davi2009,davi2011,onarlioglu2010,gupta2013,pappas2012,lu2011,lu2011a,homescu2012,gokta2014,cheng2014,pappas2013,snow2013,hiser2012,wartell2012} or detecting ROP
+attacks\cite{polychronakis2011}. Metastudies have looked at the
+effectiveness of these defenses~\cite{skowyra2013,schuster2014}.
+
+Finally, other research has shown that even with perfect control-flow
+integrity, due to the impreciseness of statically creating the
+control-flow graph, an attacker can still hijack the control
+flow~\cite{carlini2015,schuster2015,carlini2014,davi2014}.
+
+\subsubsection{Logic and implementation flaws}
+
+Not all security vulnerabilities are the result of memory corruption
+or other programming language-based flaws. Another kind of
+vulnerability class is called logic (or implementation) flaws. These
+flaws occur when the logic of the application is faulty. Consider an
+e-commerce application that allows a coupon to be reapplied to an order
+multiple times, driving the price to zero. Or if a user is able to
+enter a negative quantity for the purchase of an item, and the
+application uses that negative quantity to refund the purchase price!
+
+Logic vulnerabilities are particularly difficult for automated
+vulnerability analysis because they are deviations from the
+system's specification and the system's implementation. Very often, a
+system's specification exists only in the minds of the developers.
+Therefore, logic flaws are specific to every application, and are not
+simply the result of faulty data flows as most other vulnerability
+classes are.
+
+Research in logic flaws has typically focused on detecting a certain
+type of logic flaw, rather than detecting logic flaws generally. Early
+work looked at detecting state violations in web
+applications~\cite{cova2007}.
+
+Other research has looked at preventing access control
+vulnerabilities~\cite{dalton09:nemesis}, or statically detecting
+them~\cite{sun2011}.
+Some research has looked at other types of logic
+flaws, such as
+Privilege escalation vulnerabilities~\cite{monshizadeh2014}, data disclosure
+vulnerabilities~\cite{muthukumaran2015}, single sign-on vulnerabilities~\cite{zhou2014}, and parameter tampering vulnerabilities~\cite{bisht2010,bisht2011}.
+
+Execution After Redirect is another type of logic flaw
+vulnerability in web applications~\cite{doupe2011,payet2013}.
+
+Work on detecting logic vulnerabilities directly started with
+attempting to infer security specifications of an
+application~\cite{tan08:autoises,Livshits2009a}. Other approaches dynamically
+execute the application to infer likely invariants, then use static
+analysis to attempt to find paths that violate those
+invariants~\cite{felmetsger10:logic}. Further work tried to find missing
+security (authorization) checks automatically, without prior knowledge
+of the checks~\cite{son2011,son2013}.
+
+% citation needed here for black-box testing
+Recent work has looked at detecting logic flaw vulnerabilities in web
+applications in a black-box manner---that is, with no knowledge of the
+application's source code~\cite{pellegrino2014}.
+
+\subsubsection{Mitigation and recovery}
+%% Additions and changes made.
+%%Changes made after grammerly -- thanks Sai..its superb tool
+Mitigation and recovery are of major concern when it comes to application and systems security.
+This area of research is important because recovery of a tainted system is the first concern post-attack or exploitation.
+
+Researchers have found some Virtual Machine based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
+They try to achieve recovery in three phases, Routine execution phase, Damage assessment phase and Recovery phase.
+This framework tries to log pre-attack states and system calls and takes forward or reverse taint-tracking approach.
+Recovery phase tries to Quarantine the malicious files and objects identified during the Damage assessment.
+
+A similar VM-based approach is taken in~\cite{oliveria2008b}.
+The Bezoar framework proposed in this paper recovers a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
+The framework replays the checkpointed run while ignoring inputs from malicious sources.
+Checkpointing methodology is also used in ExecRecorder[cite], which works on the idea of log-based recovery.
+They propose a method to checkpoint system states and logging nondeterministic events.
+Recovery of the system is achieved by replaying the execution post attack and based on the log generated the system mitigates the attack execution during the reply.
+
+Recovering programs from software bugs is also a method proposed for mitigation of software failures.
+Rx~\cite{qin2007} thoroughly explains a method based on checkpointing to detect and recover from software bugs.
+This framework detects software bugs in a program using sensors and once the bug is detected mitigation is done by changing the execution environment dynamically.
+They have tabulated results displaying the mitigation of software bugs such as buffer overflow and stack overflow.
+
+Automated intrusion recovery system is proposed in~\cite{goel2005b}.
+Researchers have designed automated resolution policies which determine set of tainted files.
+Taser recovery system~\cite{goel2005b} is designed based on following models.
+Recovery model, Auditor, Analyzer and Resolver.
+Techniques like naming conventions and file permissions, logging of all events, the dependency of files, process and sockets are used by Taser in order to analyze intrusion.
+Similar to Recovery phase approach taken in~\cite{zhang2009b}, Taser uses the Resolver model proposed to go back to the state of the operating system before the intrusion.
+
+\subsubsection{Modeling and monitoring execution}
+This subarea of application layer security concentrates on modeling, monitoring, and analyzing the execution of an application to verify that the system is executing as expected, and to detect any malicious behavior.
+This can be seen as a testing strategy toward application security.
+
+There has been some research in finding new approaches of test generation for systems.
+One such approach is Concolic Testing~\cite{sen2007concolic}, which combines symbolic execution with concrete execution of the code being tested.
+
+Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{garg2013,jayaraman2009,sen2005}.
+Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
+There has also been some work on discussing concolic testing tools and their limitations~\cite{qu2011}.
+
+There has been significant research in the area of analysis of an application.
+Dynamic analysis~\cite{ball1999concept} --- one such approach --- analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{chang2008}, Anomaly Detection~\cite{cova2007,li11:block,li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{saxena2011,samuel2011}, Binary Analysis~\cite{song2008}, and Logic vulnerabilities~\cite{felmetsger2010}.
+Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{younan2012}.
+
+The other approach to analyze applications is Static Analysis~\cite{cousot1977abstract}, where the code is analyzed at compile time. 
+There has been considerable research in this area, primarily focusing on analyzing input validation and sanitization~\cite{scholte2012,yu2011,bek,balzarotti2008,balzarotti07:multi_module_va}, access control~\cite{sun2011,naumovich2004}, automatically detecting and fixing vulnerabilities in major languages~\cite{xie2006,livshits:java_static_analysis,huang:webapp_stat_analysis}, and detecting XSS and SQL injection vulnerabilities~\cite{jovanovic2006a,jovanovic06:pixy-short,bisht2008,halfond:amnesia,wassermann2004}.
+
+While a majority of research in this area is concentrated on popular attacks like XSS and SQL Injection, there is also some research on other topics like side-channel leaks~\cite{zhang2010} and information flow analysis~\cite{liu2010static,tripp2009,Livshits2009a}.
+
+\subsubsection{Protection against malicious code}
+One of the more challenging areas of application security is allowing arbitrary code execution without compromising security.
+There are situations in which applications need to run user provided code, reuse code that was written in a non-type-safe language, or host another program.
+For instance, browsers running web applications, or apps being uploaded to app stores.
+To run code safely and make sure that they are not executing malicious code, there are a number of protection mechanisms that can be used, one of which is Sandboxing, also known as Software Fault Isolation (SFI)~\cite{wahbe1993}.
+The idea is to run the code in an isolated environment within the application's address space (called fault domain) and then prevent the code from accessing addresses outside that fault domain.
+
+Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{wahbe1993,mccamant2006}.
+Native Client (NaCl)~\cite{yee2009} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
+Therefore, web developers can leverage the performance of executing native code in a safe environment.
+Xax~\cite{douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
+Other research extends this technology by covering ARM and x86-64~\cite{zhao2011,sehr2010,ansel}.
+Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{ford2008,jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{sun2014,afonso2016}.
+
+The other approach toward protecting against malicious code is to analyze the code and detect malicious behavior~\cite{zhou2012}.
+Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer~\cite{lockheimer:2012} using different techniques.
+Bouncer uses sandboxing techniques to run apps in a confined environment, but researchers~\cite{oberheide2012dissecting} have shown that apps can fingerprint Bouncer and the environment in which they are being analyzed, using evasion and obfuscating techniques to complicate the reverse engineering process.
+If apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{maier2014,jing2014,vidas2014}.
+Other approaches use analysis to detect logic bombs, i.e., when an application changes its behavior under certain conditions~\cite{fratantonio2016}.
+
+\subsection{Focus Areas}
+
+Much of the prior research focuses on just looking at memory
+corruption bugs, but we need to look at how the same sort of advanced
+protections can be applied to other classes of bugs, i.e., what is
+next? This includes: input validation, logic problems, leaking
+secrets, performance degradation, denial-of-service, command (e.g.,
+SQL) injection, and concurrency attacks.
+
+It is clear that an large amount of research effort has looked at
+securing the application layer. However, even so, there are areas that
+we believe are underserved by the research community and provide a
+significant opportunity to solve.
+
+\begin{enumerate}
+	\item The next generation of vulnerabilities (e.g., logic flaws)
+	\item Need good tools for analysis of race conditions / concurrency attacks
+\end{enumerate}
+
+This focus of this layer includes the runtime environments and abstractions
+below a high-level programming language that convert it into bytecode.
+
+\subsection{Focus Areas}
+
+\begin{enumerate}
+	\item Dealing with bloat and reducing the attack surface
+	\item Instrumentation is easier at this layer, but faster ways to observe execution are needed
+	\item Controlling easily misued middleware APIs (confused deputy)
+\end{enumerate}
 
@@
 \begin{acks}
-The authors would like to thank Dr. Maura Turolla of Telecom
-Italia for providing specifications about the application scenario.
+
 \end{acks}

### 2016-08-23 paper.tex
Finished adding Application layer, added a new citation for static info flow analysis

@@
 Attempts have been made to measure an application's attack
-surface~\cite{howard2003,manadhata2007,manadhata2008,ruprecht2014}.
+surface~\cite{howard2003, manadhata2007, manadhata2008, ruprecht2014}.
 This has the benefit of attempting to quantify the attack surface of
@@
 attempt to leverage compile-time configurability to reduce the attack
-surface~\cite{tartler2012,kurmus2013,ruprecht2014}. A different
+surface~\cite{tartler2012, kurmus2013, ruprecht2014}. A different
 approach is to reduce the attack surface by partitioning the
-application into different privilege levels~\cite{barth2010,xu2012}.
+application into different privilege levels~\cite{barth2010, xu2012}.
 Other approaches attempt to restrict the communication between
@@
 that the control-flow of the application does not
-deviate~\cite{abadi2005,abadi2009}. The first implementations were
+deviate~\cite{abadi2005, abadi2009}. The first implementations were
 deemed infeasible, so later research attempted to improve the
-efficiency of control-flow integrity~\cite{zhang2013,tice2014,bletsch2011,zhang2013,zhang2013a}.
+efficiency of control-flow integrity~\cite{zhang2013, tice2014,
+	bletsch2011, zhang2013, zhang2013a}.
 
@@
 application, so that the attacker can hijack the control-flow of the
-application without introducing new code~\cite{roemer2012,bletsch2010,tran2011}. 
-Researchers have also improved these ROP
-techniques~\cite{checkoway2010,buchanan2008,kayaalp2012,chen2012,goktas2014,schwartz2011}
- and applied them to a different
+application without introducing new code~\cite{roemer2012,
+	bletsch2010, tran2011}. Researchers have also improved these ROP
+techniques~\cite{checkoway2010, buchanan2008, kayaalp2012, chen2012,
+	goktas2014, schwartz2011} and applied them to a different
 architecture~\cite{davi2012}.
 
-Other research has looked at defending ROP attacks~\cite{pappas2012,davi2009,davi2011,onarlioglu2010,gupta2013,pappas2012,lu2011,lu2011a,homescu2012,gokta2014,cheng2014,pappas2013,snow2013,hiser2012,wartell2012} or detecting ROP
+Other research has looked at defending ROP attacks~\cite{pappas2012,
+	davi2009, davi2011, onarlioglu2010, gupta2013, pappas2012, lu2011,
+	lu2011a, homescu2012, gokta2014, cheng2014, pappas2013, snow2013,
+	hiser2012, wartell2012} or detecting ROP
 attacks\cite{polychronakis2011}. Metastudies have looked at the
-effectiveness of these defenses~\cite{skowyra2013,schuster2014}.
+effectiveness of these defenses~\cite{skowyra2013, schuster2014}.
 
@@
 control-flow graph, an attacker can still hijack the control
-flow~\cite{carlini2015,schuster2015,carlini2014,davi2014}.
+flow~\cite{carlini2015, schuster2015, carlini2014, davi2014}.
 
@@
 Privilege escalation vulnerabilities~\cite{monshizadeh2014}, data disclosure
-vulnerabilities~\cite{muthukumaran2015}, single sign-on vulnerabilities~\cite{zhou2014}, and parameter tampering vulnerabilities~\cite{bisht2010,bisht2011}.
+vulnerabilities~\cite{muthukumaran2015}, single sign-on vulnerabilities~\cite{zhou2014}, and parameter tampering vulnerabilities~\cite{bisht2010, bisht2011}.
 
 Execution After Redirect is another type of logic flaw
-vulnerability in web applications~\cite{doupe2011,payet2013}.
+vulnerability in web applications~\cite{doupe2011, payet2013}.
 
@@
 attempting to infer security specifications of an
-application~\cite{tan08:autoises,Livshits2009a}. Other approaches dynamically
+application~\cite{tan08:autoises, Livshits2009a}. Other approaches dynamically
 execute the application to infer likely invariants, then use static
@@
 security (authorization) checks automatically, without prior knowledge
-of the checks~\cite{son2011,son2013}.
+of the checks~\cite{son2011, son2013}.
 
@@
 
-Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{garg2013,jayaraman2009,sen2005}.
+Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{garg2013, jayaraman2009, sen2005}.
 Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
@@
 There has been significant research in the area of analysis of an application.
-Dynamic analysis~\cite{ball1999concept} --- one such approach --- analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{chang2008}, Anomaly Detection~\cite{cova2007,li11:block,li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{saxena2011,samuel2011}, Binary Analysis~\cite{song2008}, and Logic vulnerabilities~\cite{felmetsger2010}.
+Dynamic analysis~\cite{ball1999concept} --- one such approach --- analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{chang2008}, Anomaly Detection~\cite{cova2007, li11:block, li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{saxena2011, samuel2011}, Binary Analysis~\cite{song2008}, and Logic vulnerabilities~\cite{felmetsger2010}.
 Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{younan2012}.
@@
 
-While a majority of research in this area is concentrated on popular attacks like XSS and SQL Injection, there is also some research on other topics like side-channel leaks~\cite{zhang2010} and information flow analysis~\cite{liu2010static,tripp2009,Livshits2009a}.
+While a majority of research in this area is concentrated on popular attacks like XSS and SQL Injection, there is also some research on other topics like side-channel leaks~\cite{zhang2010} and information flow analysis~\cite{liu2010static,tripp2009,livshits2009a}.
 
@@
 
-Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{wahbe1993,mccamant2006}.
+Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{wahbe1993, mccamant2006}.
 Native Client (NaCl)~\cite{yee2009} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
@@
 Xax~\cite{douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
-Other research extends this technology by covering ARM and x86-64~\cite{zhao2011,sehr2010,ansel}.
-Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{ford2008,jana2011}.
-Sandboxing is also used for executing untrusted native code in Android applications~\cite{sun2014,afonso2016}.
+Other research extends this technology by covering ARM and x86-64~\cite{zhao2011, sehr2010, ansel}.
+Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{ford2008, jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{sun2014, afonso2016}.
 
@@
 If apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
-A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{maier2014,jing2014,vidas2014}.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{maier2014, jing2014, vidas2014}.
 Other approaches use analysis to detect logic bombs, i.e., when an application changes its behavior under certain conditions~\cite{fratantonio2016}.

### 2016-08-23 paper.tex
cleanup and adding more todos

--- 2016-08-22 paper.tex
@@
 
-%TODO Adam: Not sure if anon submission so just leaving like this for now
+% TODO Adam: Not sure if anon submission so just leaving like this for now
 \author{
@@
 \begin{abstract}
-Multifrequency media access control has been well understood in
-general wireless ad hoc networks, while in wireless sensor networks,
-researchers still focus on single frequency solutions. In wireless
-sensor networks, each device is typically equipped with a single
-radio transceiver and applications adopt much smaller packet sizes
-compared to those in general wireless ad hoc networks. Hence, the
-multifrequency MAC protocols proposed for general wireless ad hoc
-networks are not suitable for wireless sensor network applications,
-which we further demonstrate through our simulation experiments. In
-this article, we propose MMSN, which takes advantage of
-multifrequency availability while, at the same time, takes into
-consideration the restrictions of wireless sensor networks. Through
-extensive experiments, MMSN exhibits the prominent ability to utilize
-parallel transmissions among neighboring nodes. When multiple physical
-frequencies are available, it also achieves increased energy
-efficiency, demonstrating the ability to work against radio
-interference and the tolerance to a wide range of measured time
-synchronization errors.
+Abstract goes here
 \end{abstract}
@@
 
-%TODO Sai: Add keywords here
+% TODO Sai: Add keywords here
 \keywords{}
 
-\acmformat{, 2016. Total Platform Cyber Protection.}
+\acmformat{\emph{Data needed here}, 2016. Total Platform Cyber Protection.}
 % At a minimum you need to supply the author names, year and a title.
@@
 
+% TODO Sai: make `Suggested Focus Areas` and `Focus Areas` consistent 
 \subsection{Suggested Focus Areas}
@@
 \subsection{Focus Areas}
-
+% TODO Sai: Definitely needs a smoother transition from above, why are we saying research here 
+%%is inadequate and looks at only mem-corruption bugs? what is the basis?
 Much of the prior research focuses on just looking at memory
@@
 
-\subsection{Focus Areas}
-
-\begin{enumerate}
-	\item Dealing with bloat and reducing the attack surface
-	\item Instrumentation is easier at this layer, but faster ways to observe execution are needed
-	\item Controlling easily misued middleware APIs (confused deputy)
-\end{enumerate}
+% TODO Sai: WHAT is this?
+%\subsection{Focus Areas}
+%
+%\begin{enumerate}
+%	\item Dealing with bloat and reducing the attack surface
+%	\item Instrumentation is easier at this layer, but faster ways to observe execution are needed
+%	\item Controlling easily misused middleware APIs (confused deputy)
+%\end{enumerate}
 
@@
 
-In this article, we develop the first multifrequency MAC protocol for
-WSN applications in which each device adopts a
-single radio transceiver. The different MAC design requirements for
-WSNs and general wireless ad-hoc networks are
-compared, and a complete WSN multifrequency MAC design (MMSN) is
-put forth. During the MMSN design, we analyze and evaluate different
-choices for frequency assignments and also discuss the nonuniform
-back-off algorithms for the slotted media access design.
-
-% Appendix
-\appendix
-\section*{APPENDIX}
-\setcounter{section}{1}
-In this appendix, we measure
-the channel switching time of Micaz [CROSSBOW] sensor devices.
-In our experiments, one mote alternatingly switches between Channels
-11 and 12. Every time after the node switches to a channel, it sends
-out a packet immediately and then changes to a new channel as soon
-as the transmission is finished. We measure the
-number of packets the test mote can send in 10 seconds, denoted as
-$N_{1}$. In contrast, we also measure the same value of the test
-mote without switching channels, denoted as $N_{2}$. We calculate
-the channel-switching time $s$ as
-\begin{eqnarray}%
-s=\frac{10}{N_{1}}-\frac{10}{N_{2}}. \nonumber
-\end{eqnarray}%
-By repeating the experiments 100 times, we get the average
-channel-switching time of Micaz motes: 24.3$\mu$s.
-
-\appendixhead{ZHOU}
-
-% Acknowledgments
-\begin{acks}
-
-\end{acks}
+%% Appendix
+%\appendix
+%\section*{APPENDIX}
+%\setcounter{section}{1}
+%\appendixhead{}
+%
+%% Acknowledgments
+%\begin{acks}
+%
+%\end{acks}
 

### 2016-08-23 paper.tex
removed metadata stuff and added dummy dates prediction oct 2016 ;)

@@
 % Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2010}
-\acmMonth{3}
+%\acmVolume{9}
+%\acmNumber{4}
+\acmArticle{37}
+\acmYear{2016}
+\acmMonth{10}
 
@@
 \begin{bottomstuff}
-This work is supported by the National Science Foundation, under
-grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-
-Author's addresses: G. Zhou, Computer Science Department,
-College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-Computer Science Department, University of Virginia; T. Yan,
-Eaton Innovation Center; T. He, Computer Science Department,
-University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-(Current address) NASA Ames Research Center, Moffett Field, California 94035.
+	Need to add stuff here
 \end{bottomstuff}
@@
 
+% TODO Sai: What features??
 These features of Trojan logic in hardware drive the research that
@@
 % History dates
-\received{February 2007}{March 2009}{June 2009}
+\received{September 2016}{October 2016}{October 2016}
 
 \end{document}
-% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM
